Finiteness and Cardinality in Homotopy Type Theory

### 2020-01-08 paper.tex
more

  It is defined as a Higher Inductive Type:
  \begin{equation}
    \begin{aligned}
      \lVert A \rVert &= \\
        &| \; \lvert \cdot \rvert : A \rightarrow \lVert A \rVert ; \\
        &| \; \text{squash} : \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; \\
    \end{aligned}
  \end{equation}

@@

      \mathcal{K}(A) &=  \\
        &\left\vert \; \cdot \dblcolon \cdot : A \times \mathcal{K}(A) \rightarrow \mathcal{K}(A) \right; \\
        &\left\vert \; [] : \mathcal{K}(A) \right; \\
        &\left\vert \; \text{com} : \Pi (x, y: A) , \Pi (\mathit{xs}: \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} \right; \\
        &\left\vert \; \text{dup} : \Pi (\mathit{x}: A) , \Pi (\mathit{xs}: \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} \right; \\
        &\left\vert \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q \right;

@@

Other representations of \(\mathcal{K}\) \cite{fruminFiniteSetsHomotopy2018}
are more explicit constructions of the free join semilattice (i.e. there is a
point constructors for union instead of cons, and then path constructors for
the associativity and identity laws), but we have found this representation
easier to work with.
Nonetheless, the alternative representation is included in our formalisation,
and proven equivalent to the representation here.
\begin{rm-definition}[Membership of \mathcal{K}]
  First, we need to provide equations for the two point constructors.
  \begin{equation}
    \begin{alignat}{2}
      x &\in& [] &= \bot \\
      x &\in& y \dblcolon \mathit{xs} &= \lVert (x \equiv y) + (x \in \mathit{xs}) \rVert
    \end{alignat}
  \end{equation}
  The \(\text{com}\) and \(\text{dup}\) constructors are handled by proving that
  the truncated form of \(+\) is itself commutative and idempotent.
  The type of propositions is itself a set, satisfying the \(\text{trunc}\)
  constructor.
\end{rm-definition}
\begin{rm-definition}[Kuratowski Finiteness]
  A type is Kuratowski finite iff there exists a Kuratowski Set which contains
  all of its elements.
  \begin{equation}
    \mathcal{K}^{f}(A) = \Sigma {(\mathit{xs} : \mathcal{K}(A))} , \Pi (x : A) , x \in \mathit{xs}
  \end{equation}
\end{rm-definition}

\subsection{Strength}
Since the Kuratowski set is a departure in structure from our previous
list-based notions of finiteness, it makes sense to first look for the closest
list-based analogue.
As it turns out, that analogue is manifestly enumerable finiteness, with the
order removed.
\begin{rm-theorem}
  A proof of Kuratowski finiteness is equivalent to a propositionally truncated
  proof of enumerability.
  \begin{equation}
    \mathcal{K}^f(A) \simeq \lVert \mathcal{E}(A) \rVert
  \end{equation}
\end{rm-theorem}
\begin{proof}
  We prove by way of an isomorphism.
  In the first direction (from \(\mathcal{K}\) to \(\mathcal{E}\)), because we
  are eliminating into a proposition, we need only deal with the point
  constructors.
  For these, we convert the \(\mathcal{K}\) cons to its list counterpart, and
  similarly for the nil constructor.

@@

\subsection{Closure}

>>>>

  It is defined as a Higher Inductive Type:
  \begin{equation}
    \begin{aligned}
      \lVert A \rVert =& \; \lvert \cdot \rvert : A \rightarrow \lVert A \rVert ; \\
                      |& \; \text{squash} : \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; 
    \end{aligned}
  \end{equation}

@@

      \mathcal{K}(A) =&
       \; \cdot \dblcolon \cdot : A \times \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
      \vert& \; [] : \mathcal{K}(A) ; \\
      \vert& \; \text{com} : \Pi (x, y: A) , \Pi (\mathit{xs}: \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
      \vert& \; \text{dup} : \Pi (\mathit{x}: A) , \Pi (\mathit{xs}: \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} ; \\
      \vert& \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;

@@

Other representations of \(\mathcal{K}\) \cite{fruminFiniteSetsHomotopy2018}
are more explicit constructions of the free join semilattice (i.e. there is a
point constructors for union instead of cons, and then path constructors for
the associativity and identity laws), but we have found this representation
easier to work with.
Nonetheless, the alternative representation is included in our formalisation,
and proven equivalent to the representation here.
\begin{rm-definition}[Membership of \mathcal{K}]
  First, we need to provide equations for the two point constructors.
  \(x \in [] = \bot\), and \(x \in y \dblcolon \mathit{xs} = \lVert (x \equiv y)
  + (x \in \mathit{xs}) \rVert \).
  The \(\text{com}\) and \(\text{dup}\) constructors are handled by proving that
  the truncated form of \(+\) is itself commutative and idempotent.
  The type of propositions is itself a set, satisfying the \(\text{trunc}\)
  constructor.
\end{rm-definition}
\begin{rm-definition}[Kuratowski Finiteness]
  A type is Kuratowski finite iff there exists a Kuratowski Set which contains
  all of its elements.
  \begin{equation}
    \mathcal{K}^{f}(A) = \Sigma {(\mathit{xs} : \mathcal{K}(A))} , \Pi (x : A) , x \in \mathit{xs}
  \end{equation}
\end{rm-definition}

\begin{rm-theorem}
  A proof of Kuratowski finiteness is equivalent to a propositionally truncated
  proof of enumerability.
  \begin{equation}
    \mathcal{K}^f(A) \simeq \lVert \mathcal{E}(A) \rVert
  \end{equation}
\end{rm-theorem}
\begin{proof}
  We prove by way of an isomorphism.
  In the first direction (from \(\mathcal{K}\) to \(\mathcal{E}\)), because we
  are eliminating into a proposition, we need only deal with the point
  constructors.
  For these, we convert the \(\mathcal{K}\) cons to its list counterpart, and
  similarly for the nil constructor.

@@
